---
title: "Fine mapping analysis of suggestive sentinel variants for severe asthma study"
author: "Noemi Nicole Piga"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: "hide"
    toc: true
    toc_float: true
    theme: united
---

```{r, setup, hide = TRUE}
#library reticulate to be able to add chunks in another language
#library(reticulate)
#To run this .rmd file in a terminal being in the project folder:
#export PATH=${PATH}:/cm/shared/apps/R/deps/rstudio/bin/pandoc
#file="./report/Fine_mapping_suggestive_sentinel_variants.Rmd"
#module unload R/4.2.1
#module load R/4.1.0
#Rscript -e 'rmarkdown::render("'$file'")'
```


# Rationale
Fine mapping of replicated suggestive sentinel variants discovered in the severe asthma GWAS on UKBiobank European dataset and replicated by meta-analysis of other three moderate-to-severe and severe cohorts. I used two different tools and compare the credible sets.

# Analysis
Code: src/bgenix_index.sh
```{r, code = readLines('/home/n/nnp5/PhD/PhD_project/Fine_mapping_severe_asthma/src/bgenix_index.sh'), eval=FALSE}
```
Code: src/finemapping_replicated_suggestive/FINEMAP_replicated_suggestive.sh
```{r, code = readLines('/home/n/nnp5/PhD/PhD_project/Fine_mapping_severe_asthma/src/finemapping_replicated_suggestive/FINEMAP_replicated_suggestive.sh'), eval=FALSE}
```
Code: src/finemapping_replicated_suggestive/credset_FINEMAP.R
```{r, code = readLines('/home/n/nnp5/PhD/PhD_project/Fine_mapping_severe_asthma/src/finemapping_replicated_suggestive/credset_FINEMAP.R'), eval=FALSE}
```
Code: src/finemapping_replicated_suggestive/Susie_replicated_suggestive.sh
```{r, code = readLines('/home/n/nnp5/PhD/PhD_project/Fine_mapping_severe_asthma/src/finemapping_replicated_suggestive/Susie_replicated_suggestive.sh'), eval=FALSE}
```
Code: src/z_score.R
```{r, code = readLines('/home/n/nnp5/PhD/PhD_project/Fine_mapping_severe_asthma/src/z_score.R'), eval=FALSE}
```
Code: src/finemapping_replicated_suggestive/Susie_replicated_suggestive.R
```{r, code = readLines('/home/n/nnp5/PhD/PhD_project/Fine_mapping_severe_asthma/src/finemapping_replicated_suggestive/Susie_replicated_suggestive.R'), eval=FALSE}
```
Code: src/finemapping_replicated_suggestive/credset_susie.R
```{r, code = readLines('/home/n/nnp5/PhD/PhD_project/Fine_mapping_severe_asthma/src/finemapping_replicated_suggestive/credset_susie.R'), eval=FALSE}
```

Two different fine-mapping tools were implemented, SuSiE [1] and FINEMAP [2].
<br>
For both tools, I investigated the genomic regions of 1Mb centred on the replicated variant, I allowed up to 10 causal variants, and I tested for 46,086 individuals with in-sample LD matrix created with LDstore2 [4] for FINEMAP, plink2 [5] and R for SuSiE. If a software found multiple credible sets for one region, I included all of them. If a software failed, I considered as valid the credible set found by the other tool. 
<br>
To assess the final number of variants to follow-up, I followed an approach as taken by Kanai M. et al. [6]: consider the credible set variants found in both tools; keep variants whose SuSiE's PIP and FINEMAP's PIP differ less than 0.05; calculate an average PIP for each variant. Kanai M. et al., found that variants with a consistent PIP difference (greater than 0.05) denoted poor enirchment ffor known functional annotations and more often resulted in false positives [6].

# Results


# Conclusion
Two different tools, SuSiE and FINEMAP, were used to fine-map 21 variants. After validation of the credible set I obtained XXX

# References
1.Zou Y. et al., Fine-mapping from summary data with the "Sum of Single Effects" model. PLoS Genet. 2022 Jul 19;18(7):e1010299. doi: 10.1371/journal.pgen.1010299. PMID: 35853082; PMCID: PMC9337707.
2.Benner C. et al., FINEMAP: efficient variable selection using summary data from genome-wide association studies. Bioinformatics. 2016 May 15;32(10):1493-501. doi: 10.1093/bioinformatics/btw018. Epub 2016 Jan 14. PMID: 26773131; PMCID: PMC4866522.
3.Weissbrod, O. et al. Functionally informed fine-mapping and polygenic localization of complex trait heritability. Nat Genet 52, 1355–1363 (2020). https://doi.org/10.1038/s41588-020-00735-5.
4.Benner, C. et al. Prospects of fine-papping trait-associated genomic regions by using summary statistics from genome-wide association studies. Am. J. Hum. Genet. (2017).
5.Christopher C Chang et al., Second-generation PLINK: rising to the challenge of larger and richer datasets, GigaScience, December 2015, doi:https://doi.org/10.1186/s13742-015-0047-8
6.Kanai M. et al., Insights from complex trait fine-mapping across diverse populations, medRxiv 2021.09.03.21262975; doi: https://doi.org/10.1101/2021.09.03.21262975.

# Notes

