---
title: "Fine mapping analysis of suggestive sentinel variants for severe asthma study"
author: "Noemi Nicole Piga"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: "hide"
    toc: true
    toc_float: true
    theme: united
---

```{r, setup, hide = TRUE}
#library reticulate to be able to add chunks in another language
#library(reticulate)
#To run this .rmd file in a terminal being in the project folder:
#export PATH=${PATH}:/cm/shared/apps/R/deps/rstudio/bin/pandoc
#file="./report/Fine_mapping_suggestive_sentinel_variants.Rmd"
#Rscript -e 'rmarkdown::render("'$file'")'
```


# Rationale
Fine mapping of replicated suggestive sentinel variants discovered in the severe asthma GWAS on UKBiobank European dataset and replicated by meta-analysis of other three moderate-to-severe and severe cohorts. I used two different tools and compare the credible sets.

# Analysis
Code: src/bgenix_index.sh
```{r, code = readLines('/home/n/nnp5/PhD/PhD_project/Fine_mapping_severe_asthma/src/bgenix_index.sh'), eval=FALSE}
```
Code: src/finemapping_replicated_suggestive/Merge_loci_for_finemapping.sh
```{r, code = readLines('/home/n/nnp5/PhD/PhD_project/Fine_mapping_severe_asthma/src/finemapping_replicated_suggestive/Merge_loci_for_finemapping.sh'), eval=FALSE}
```
Code: src/finemapping_replicated_suggestive/Merge_loci_for_finemapping.R
```{r, code = readLines('/home/n/nnp5/PhD/PhD_project/Fine_mapping_severe_asthma/src/finemapping_replicated_suggestive/Merge_loci_for_finemapping.R'), eval=FALSE}
```
Code: src/finemapping_replicated_suggestive/FINEMAP_replicated_suggestive.sh
```{r, code = readLines('/home/n/nnp5/PhD/PhD_project/Fine_mapping_severe_asthma/src/finemapping_replicated_suggestive/FINEMAP_replicated_suggestive.sh'), eval=FALSE}
```
Code: src/finemapping_replicated_suggestive/credset_FINEMAP.R
```{r, code = readLines('/home/n/nnp5/PhD/PhD_project/Fine_mapping_severe_asthma/src/finemapping_replicated_suggestive/credset_FINEMAP.R'), eval=FALSE}
```
Code: src/finemapping_replicated_suggestive/Susie_replicated_suggestive.sh
```{r, code = readLines('/home/n/nnp5/PhD/PhD_project/Fine_mapping_severe_asthma/src/finemapping_replicated_suggestive/Susie_replicated_suggestive.sh'), eval=FALSE}
```
Code: src/z_score.R
```{r, code = readLines('/home/n/nnp5/PhD/PhD_project/Fine_mapping_severe_asthma/src/z_score.R'), eval=FALSE}
```
Code: src/finemapping_replicated_suggestive/Susie_replicated_suggestive.R
```{r, code = readLines('/home/n/nnp5/PhD/PhD_project/Fine_mapping_severe_asthma/src/finemapping_replicated_suggestive/Susie_replicated_suggestive.R'), eval=FALSE}
```
Code: src/finemapping_replicated_suggestive/credset_susie.R
```{r, code = readLines('/home/n/nnp5/PhD/PhD_project/Fine_mapping_severe_asthma/src/finemapping_replicated_suggestive/credset_susie.R'), eval=FALSE}
```
Code: src/finemapping_replicated_suggestive/finemapping_visualisation.sh
```{r, code = readLines('/home/n/nnp5/PhD/PhD_project/Fine_mapping_severe_asthma/src/finemapping_replicated_suggestive/finemapping_visualisation.sh'), eval=FALSE}
```
Code: src/finemapping_replicated_suggestive/PIP_plot_with_funcannot.R
```{r, code = readLines('/home/n/nnp5/PhD/PhD_project/Fine_mapping_severe_asthma/src/finemapping_replicated_suggestive/PIP_plot_with_funcannot.R'), eval=FALSE}
```
Code: src/finemapping_replicated_suggestive/PIP_plot_with_funcannot_pt2.R
```{r, code = readLines('/home/n/nnp5/PhD/PhD_project/Fine_mapping_severe_asthma/src/finemapping_replicated_suggestive/PIP_plot_with_funcannot_pt2.R'), eval=FALSE}
```


Two different fine-mapping tools were implemented, SuSiE-RSS [1] and FINEMAP [2]. SuSiE-RSS (stands for 'Sum of Single Effects - Regression with summary statistics') runs finemapping analysis using summary statistics and z-score as sufficient statistics; it uses Bayesian Stepwise Selection algorithm [3] that iteratively tests for each possible number of multiple causal varaints in the locus, up to ten. FINEMAP uses summary statistics as well and it implements a shotgun stochastic search algorithm allowing the investigation of multiple causal variant in one locus up to ten as well.
<br>
Genomic loci were defined as a region of 1Mb centred on the replicated suggestive variants; if loci overlapped, they were merged together and analysed as a unique region. For both tools, I allowed up to 10 causal variants, and I tested for 46,086 individuals with in-sample LD matrix created with LDstore2 [4] for FINEMAP, plink2 [5] and R for SuSiE. If a software found multiple credible sets for one region, I included all of them. If a software failed, I considered as valid the credible set found by the other tool.
<br>

<br>
To assess the final number of variants to follow-up, I followed an approach as taken by Kanai M. et al. [6]: consider the credible set variants found in both tools; keep variants whose SuSiE's PIP and FINEMAP's PIP differ less than 0.05; calculate an average PIP for each variant. Kanai M. et al. found that variants with a consistent PIP difference (greater than 0.05) denoted poor enirchment for known functional annotations and more often resulted in false positives [6].

# Results
After merging regions, I obtained 18 genomic loci that were investigated for fine-mapping. Using FINEMAP, I identified variants for all the credible sets while SuSiE identified credible set variants for thirteen loci. After validation of credible sets variants, 17 loci were identified with a total of 616 good quality variants for functional analysis.
```{r, echo=FALSE}
library(knitr)
library(readxl)
df  <-  read_excel("/home/n/nnp5/PhD/PhD_project/Fine_mapping_severe_asthma/src/report/Finemapping_repl_sugg_vars.xlsx", sheet="Credible_set_N", skip = 1, col_names = TRUE)
kable(df)
```

# Conclusion
Two different tools, SuSiE and FINEMAP, were used to fine-map the 21 replicated suggestive variants (18 merged regions). After validation of the credible sets, I obtained 615 fine-mapped variants to take forward for variant-to-gene function analyses.

# Update: Fine-mapping of replicated signal rs778801698 and surrounding genomic locus.
Code: src/bgenix_index_sbatch.sh
```{r, code = readLines('/home/n/nnp5/PhD/PhD_project/Fine_mapping_severe_asthma/src/bgenix_index_sbatch.sh'), eval=FALSE}
```
Code: src/Fine_mapping_chr3_rs778801698.sh
```{r, code = readLines('/home/n/nnp5/PhD/PhD_project/Fine_mapping_severe_asthma/src/Fine_mapping_chr3_rs778801698.sh'), eval=FALSE}
```

# Conclusion - update
We decided to exclude the fine-mapping of the HLA region, as it presents a complex LD-structure. We therefore do not include
the varaint-to-gene mapping results for this region as well.
<br>
We successfully fine-mapped 17 genomic loci. We added genomic locus on chromosome 3 with sentinel variant rs778801698; in this
locus we obtain a total of 224 in the 95% credible set variants. In this locus, the high number of putative causal variants is due to the
presence in both FINEMAP and SuSiE of four variants (rs778801698, rs2271960, rs2271961, rs6765484) with PIP >= 1% and all the other variants with PIP lower than 1%.
<br>
We obtained a total of 741 putative causal variants from 17 fine-mapped genomic loci; 232 variants have PIP >= 1%.



# References
1.Zou Y. et al., Fine-mapping from summary data with the "Sum of Single Effects" model. PLoS Genet. 2022 Jul 19;18(7):e1010299. doi: 10.1371/journal.pgen.1010299. PMID: 35853082; PMCID: PMC9337707.
<br>
2.C.Benner, Refining fine-mapping: effect sizes and regional heritability, bioRxiv 318618, doi:https://doi.org/10.1101/
<br>
3.Gao Wang et al., A Simple New Approach to Variable Selection in Regression, with Application to Genetic Fine Mapping, Journal of the Royal Statistical Society Series B: Statistical Methodology, December 2020, doi:https://doi.org/10.1111/rssb.12388
<br>
4.Benner, C. et al. Prospects of fine-papping trait-associated genomic regions by using summary statistics from genome-wide association studies. Am. J. Hum. Genet. (2017).
<br>
5.Christopher C Chang et al., Second-generation PLINK: rising to the challenge of larger and richer datasets, GigaScience, December 2015, doi:https://doi.org/10.1186/s13742-015-0047-8
<br>
6.Kanai M. et al., Insights from complex trait fine-mapping across diverse populations, medRxiv 2021.09.03.21262975; doi: https://doi.org/10.1101/2021.09.03.21262975.

