---
title: "Fine mapping analysis of sentinel variants for severe asthma study"
author: "Noemi Nicole Piga"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: "hide"
    toc: true
    toc_float: true
    theme: united
---

```{r, setup, hide = TRUE}
#library reticulate to be able to add chunks in another language
library(reticulate)
#To run this .rmd file in a terminal being in the project folder:
#export PATH=${PATH}:/cm/shared/apps/R/deps/rstudio/bin/pandoc
#file="./report/Fine_mapping.Rmd"
#module unload R/4.2.1
#module load R/4.1.0
#Rscript -e 'rmarkdown::render("'$file'")'
```


# Rationale
Fine mapping of sentinel variants discovered in the severe asthma GWAS on UK Biobank European dataset.

# Analysis
Code: Post_GWAS/src/Susie.sh
```{r, code = readLines('/home/n/nnp5/PhD/PhD_project/Fine_mapping_severe_asthma/src/Susie.sh'), eval=FALSE}
```
Code: Post_GWAS/src/susie.R
```{r, code = readLines('/home/n/nnp5/PhD/PhD_project/Fine_mapping_severe_asthma/src/susie.R'), eval=FALSE}
```
Code: Post_GWAS/src/FINEMAP.sh
```{r, code = readLines('/home/n/nnp5/PhD/PhD_project/Fine_mapping_severe_asthma/src/FINEMAP.sh'), eval=FALSE}
```
Code: Post_GWAS/src/pre_processing_Polyfun.sh
```{r, code = readLines('/home/n/nnp5/PhD/PhD_project/Fine_mapping_severe_asthma/src/pre_processing_Polyfun.sh'), eval=FALSE}
```
Code: Post_GWAS/src/pre_processing_Polyfun.R
```{r, code = readLines('/home/n/nnp5/PhD/PhD_project/Fine_mapping_severe_asthma/src/pre_processing_Polyfun.R'), eval=FALSE}
```
Code: Post_GWAS/src/PolyFun_approach3.sh
```{r, code = readLines('/home/n/nnp5/PhD/PhD_project/Fine_mapping_severe_asthma/src/PolyFun_approach3.sh'), eval=FALSE}
```

Four different fine-mapping tools were implemented, more precisely: I tested and compared the results for Susie [1] and FINEMAP [2] tools with and without functional information given as a prior for the model. For the funcitonal informed fine-mapping, I used the PolyFUN tool [3].


# Results

# Conclusion

# References


# Notes

Add image:
```{r, out.width='100%', out.height='200%', fig.align='center', fig.cap='Fig 5', echo = FALSE, eval=FALSE}
knitr::include_graphics('')
```
